\thetitle

### 2020-07-20 writing/mutate.tex
working on mutation

So far we have established that our mutators and~\citet{gtnffvf-jfp-2019}'s 
benchmarks are sufficient to create a large number of realistic ill-typed Type Racket
programs. However, for ill-typed programs if their types are correct, like
in our case, it is trivial to figure out which component contains the bug.
It is the component that doesn't type check. Thus the ill-typed mutants
do not offer an interesting setting to try out the debugging abilities of
rational programmers. That said, we can derive interesting debugging
scenarios by removing type specifications from some of the components of
the benchmarks. The resulting mix-typed programs may be well-typed and when
run may raise a type error. Section~\ref{sec:rational} explains how we can
explore the space of debugging scenarios in a principled manner. As a
final remark in this section, with the data from
figure~\ref{fig:mutators-runtime-hit}, we verify  that in each of the three semantics
we target, our mutators generate mutants that half the time give rise to at least one
interesting debugging scenario.

>>>>

So far we have established that our mutators and~\citet{gtnffvf-jfp-2019}'s 
benchmarks are sufficient to create a large number of realistic ill-typed Type Racket
programs. However, in an ill-typed program with correct types, like
in our case, it is trivial to figure out which component contains the bug.
It is the component that doesn't type check. Thus the ill-typed mutants
do not offer an interesting setting to try out the debugging abilities of
rational programmers. That said, we can derive interesting debugging
scenarios by removing type specifications from some of the components of
the benchmarks. The resulting mix-typed programs may be well-typed and when
run may raise a type error. Section~\ref{sec:rational} explains how we can
explore the space of debugging scenarios in a principled manner. As a
final remark in this section, with the data from
figure~\ref{fig:mutators-runtime-hit}, we verify  that in each of the three semantics
we target, between 25\% and
90\% of the ill-typed mutants each mutator generates give rise to at least one interesting debugging scenario.

### 2020-07-20 writing/rational.tex
working on mutation

Based on this idea, a metric for migratory paths is a concise way to
evaluate the effectiveness of a rational programmer. The components of
each program create a \emph{debugging scenario space}, a lattice like
the one introduced by~\citet{tfgnvf-popl-2016} for performance
evaluations. This space consists of variants of the program, dubbed
\emph{scenarios}, that differ in their type specifications. In the
setting of migratory typing,\footnote{The method can be adapted to
so-called true gradual typing systems~\cite{svcb-snapl-2015} in a
straightforward manner.} each scenario differs from another in which
components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is a subset of
$\set{\component}$ . These scenarios of $\system$ are ordered by the
subset relation and thus form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements.  The bottom scenario of
$\lattice{\system}$ is always $\emptyset$ and the top one is $\system$
itself. In between are all the mix-typed scenarios.

@@

In detail we answer $Q_1$ for a mutant by creating trails for each natural blame,
natural exception and natural random trail in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and determine the scenario that can detect the fault in
the program. This can happen either because Typed Racket's type checker outright
rejects the scenario or because natural Typed Racket's runtime performs a type check
that fails. In either case, each of these scenarios is the root of a trail. Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail---a trail may the top of the finite lattice---we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
roots two more times; once following the natural-exception programmer and once the
natural-random programmer. Finally, based on the collected data, we calculate the
success/failure results for the trails to determine whether natural blame is more
successful than natural exceptions and pure chance and we obtain the answer for
$Q_1$.  The process is analogues for the other questions employing the respective
modes of the rational programmer.

>>>>

Based on this idea, a metric for migratory paths is a concise way to
evaluate the effectiveness of a rational programmer. The components of
an ill-typed program, such as one of  our ill-typed mutants, 
create a \emph{debugging scenario space}, a lattice like
the one introduced by ~\citet{tfgnvf-popl-2016} for performance
evaluations. As we discuss briefly at the end of ~\ref{sec:mutate}, 
this space consists of variants of the program, dubbed
\emph{scenarios}, that differ in their type specifications. In the
setting of migratory typing,\footnote{The method can be adapted to
so-called true gradual typing systems~\cite{svcb-snapl-2015} in a
straightforward manner.} each scenario differs from another in which
components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is a subset of
$\set{\component}$ . These scenarios of $\system$ are ordered by the
subset relation and thus form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements.  The bottom scenario of
$\lattice{\system}$ is always $\emptyset$ and the top one is $\system$
itself. In between are all the mix-typed scenarios.

@@

In detail we answer $Q_1$ for an ill-typed mutant from
section~\ref{sec:mutate} by creating trails for each natural blame,
natural exception and natural random trail in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and determine the scenario that can detect the fault in
the program. This can happen either because Typed Racket's type checker outright
rejects the scenario or because natural Typed Racket's runtime performs a type check
that fails. In either case, each of these scenarios is the root of a trail. Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail---a trail may the top of the finite lattice---we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
roots two more times; once following the natural-exception programmer and once the
natural-random programmer. Finally, based on the collected data, we calculate the
success/failure results for the trails to determine whether natural blame is more
successful than natural exceptions and pure chance and we obtain the answer for
$Q_1$.  The process is analogous for the other questions employing the respective
modes of the rational programmer. Figure~\ref{fig:process} summarizes the
experimental process for one mode of the rational programmer and connects
it with the mutations from section~\ref{sec:mutation}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{./Images/process}
  \caption{The experimental process for one mode of the rational
  programmer}
  \label{fig:process}
\end{figure}
